---
title: "Consequences of Severe Weather Events in USA"
output: 
  html_document:
    keep_md: true
---

## Synopsis

In this report we aim to discover the types of severe weather events across the United States that are, on the one hand, most harmful with respect to population health and have, on the other hand, the greatest economic consequences.


## Data Processing

From the [Coursera Reproducible Research Peer Assessment 2 course web site](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2) we obtained NOAAA storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage. The events in the database start in the year 1950 and end in November 2011. In the earlier years of the database there are generally fewer events recorded, most likely due to a lack of good records. More recent years should be considered more complete.

### Reading in the storm database

We first read the database from the raw text file included in the bzip2 archive. The data is a comma separated values file were fields are delimited with the comma (,) character and missing values are coded as blank fields. Initially we also read the header data.

```{r, echo=TRUE,cache=TRUE}
## read from a connection to bzip2 file
s0 <- read.csv(bzfile(description = "repdata-data-StormData.csv.bz2"), na.strings = "")
```

After reading in the database we check the first few rows (there are `r format(nrow(s0), big.mark=",")`) rows in this dataset.

```{r, echo=TRUE}
dim(s0)
```

```{r, echo=TRUE}
head(s0[, 1:8])
```

We make sure that the column headers are properly formated for R data frames.

```{r, echo=TRUE}
cnames <- names(s0)
names(s0) <- make.names(cnames)  ## Ensure names are properly formatted
head(s0[, 1:8])
```

The columns we are interested in are the EVTYPE column which contains the types of events, the FATALITIES and INJURIES columns with their respective counts, and the PROPDMG and CROPDMG with an estimate of damage in property and crops respectively. We only take the rows with one or more of this columns greater than zero. 

```{r, echo=TRUE}
s0$EVTYPE <- toupper(s0$EVTYPE)
s1 <- s0[s0$FATALITIES>0 | s0$INJURIES>0, c("EVTYPE","FATALITIES","INJURIES")]

s2 <- s0[s0$PROPDMG>0 | s0$CROPDMG>0, c("EVTYPE","PROPDMG","CROPDMG")]
```
After that we reduce the database to `r format(nrow(s1), big.mark=",")`) rows. 


## Results

### Across the United States, types of events most harmful with respect to population health, i.e., with fatalities or injuries

```{r, echo=TRUE}
a1 <- aggregate(FATALITIES+INJURIES ~ EVTYPE, data = s1 , FUN = sum)
a1 <- a1[order(-a1$`FATALITIES + INJURIES`), ]
plot(log(a1$`FATALITIES + INJURIES`), type = "l")


```

### Across the United States,types of events with the greatest economic consequences, i.e., with property or crop damage

```{r, echo=TRUE}
a2 <- aggregate(PROPDMG+CROPDMG ~ EVTYPE, data = s2 , FUN = sum)
a2 <- a2[order(-a2$`PROPDMG + CROPDMG`), ]
plot(log(a2$`PROPDMG + CROPDMG`), type = "l")
```


